---
title: "exercise72"
author: "Lev Mazaev"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 7.2 (Applied Predictive Modeling, p. 169)

### Loading the packages
```{r, echo=TRUE, message=FALSE}
library(mlbench)
library(caret)
library(doMC)
library(earth)
registerDoMC(8)
```

### Creating the data
```{r}
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
trainingData$x <- data.frame(trainingData$x)
featurePlot(trainingData$x, trainingData$y)
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```

### Correlation check
```{r}
corrplot::corrplot(cor(trainingData$x))
findCorrelation(cor(trainingData$x), cutoff = 0.75)
```
#### There are no highly-correlated predictors

### Neural Network
```{r}
nnetGrid <- expand.grid(.decay = c(0, 0.01, .1), .size = 1:10, .bag = FALSE)
nnetModel <- train(x = trainingData$x, y = trainingData$y,
                   method = "avNNet", tuneGrid = nnetGrid, 
                   trControl = trainControl(method = "cv"),
                   preProcess = c("center", "scale"),
                   linout = TRUE, trace = TRUE,
                   MaxNWts = 10 * (ncol(trainingData$x) + 1) + 10 + 1,
                   maxit = 500)
nnetModel
plot(nnetModel)
```

#### Performance evaluation
```{r}
nnetPred <- predict(nnetModel, newdata = testData$x)
modelstats <- postResample(pred = nnetPred, obs = testData$y)
postResample(pred = nnetPred, obs = testData$y)
```

### K Nearest Neighbors
```{r}
knnModel <- train(x = trainingData$x, y = trainingData$y,
                  method = "knn", preProcess = c("center", "scale"),
                  tuneLength = 10)
knnModel
plot(knnModel)
```

#### Performance evaluation
```{r}
knnPred <- predict(knnModel, newdata = testData$x)
modelstats <- rbind(modelstats, postResample(pred = knnPred, obs = testData$y))
postResample(pred = knnPred, obs = testData$y)
```

### MARS
```{r}
marsGrid <- expand.grid(.degree = 1:2, .nprune = 2:38)
marsModel <- train(x = trainingData$x, y = trainingData$y, method = "earth", 
                   tuneGrid = marsGrid, trControl = trainControl(method = "cv"))
marsModel
plot(marsModel)
```

#### Which predictors did MARS select as the informative?
```{r}
varImp(marsModel)
```
#### Only X1-X5

#### Performance evaluation
```{r}
marsPred <- predict(marsModel, newdata = testData$x)
modelstats <- rbind(modelstats, postResample(pred = marsPred, obs = testData$y))
postResample(pred = marsPred, obs = testData$y)
```

### SVM Radial
```{r}
svmRModel <- train(x = trainingData$x, y = trainingData$y,
                  method = "svmRadial", preProcess = c("center", "scale"),
                  tuneLength = 14, trControl = trainControl(method = "cv"))
svmRModel
plot(svmRModel)
svmRModel$finalModel
```

#### Performance evaluation
```{r}
svmRPred <- predict(svmRModel, newdata = testData$x)
modelstats <- rbind(modelstats, postResample(pred = svmRPred, obs = testData$y))
postResample(pred = svmRPred, obs = testData$y)
```

## Which model performs the best?
```{r}
rownames(modelstats) <- c('NNET', 'KNN', 'MARS', 'SVM')
modelstats
```

#### MARS model gives the best performance. Also it selects only informative predictors (X1-X5)